---
title: "Barplot_Sid_Sym_C"
author: "Veronica Radice"
date: "07/05/2021"
output: html_document
---


# Barplots of Differentially expressed genes

## Coral symbionts - *Cladocopium*
### *Siderastrea siderea*

### Import libraries
```{r}
library(ggplot2)
library(scales)
library(dplyr)
library(plyr)
```

### Resources
https://stackoverflow.com/questions/38268741/geom-bar-ggplot2-stacked-grouped-bar-plot-with-positive-and-negative-values-p


### Set working directory
```{r}
setwd("~/Documents/Rprojects/postdoc\ Rprojects/ODU_postdoc_Rprojects/Paper_Ojo_gene-expression/Siderastrea/symbiont/")
```


# Effect of Destination Control


Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Sid_Symbiont_DE_genes_OC_LC_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'OC-LC'
de.genes$group <- as.factor(de.genes$group)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))

summary_DEG[1,3] <- summary_DEG[1,3]*(-1)

summary_DEG_1 <- summary_DEG

summary_DEG_1
```


###############################################################################

Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Sid_Symbiont_DE_genes_OC_RC_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'OC-RC'
de.genes$group <- as.factor(de.genes$group)
dim(de.genes)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))

summary_DEG[1,3] <- summary_DEG[1,3]*(-1)

summary_DEG_2 <- summary_DEG
summary_DEG_2
```


###############################################################################

Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Sid_Symbiont_DE_genes_LC_RC_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'LC-RC'
de.genes$group <- as.factor(de.genes$group)
dim(de.genes)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))

summary_DEG[1,3] <- summary_DEG[1,3]*(-1)

summary_DEG_3 <- summary_DEG
summary_DEG_3
```


###############################################################################

# Effect of Destination Ojo

Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Sid_Symbiont_DE_genes_OO_LO_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'OO-LO'
de.genes$group <- as.factor(de.genes$group)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))

summary_DEG[1,3] <- summary_DEG[1,3]*(-1)

summary_DEG_4 <- summary_DEG
summary_DEG_4
```

###############################################################################

Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Sid_Symbiont_DE_genes_LO_RO_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'LO-RO'
de.genes$group <- as.factor(de.genes$group)
dim(de.genes)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))

summary_DEG[1,3] <- summary_DEG[1,3]*(-1)

summary_DEG_5 <- summary_DEG
summary_DEG_5
```


###############################################################################

Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Sid_Symbiont_DE_genes_OO_RO_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'OO-RO'
de.genes$group <- as.factor(de.genes$group)
dim(de.genes)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))

summary_DEG[1,3] <- summary_DEG[1,3]*(-1)

summary_DEG_6 <- summary_DEG
summary_DEG_6
```


###############################################################################

# Ojo vs Control (same origins)

Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Sid_Symbiont_DE_genes_LO_LC_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'LO-LC'
de.genes$group <- as.factor(de.genes$group)
dim(de.genes)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))

summary_DEG[1,3] <- summary_DEG[1,3]*(-1)

summary_DEG_7 <- summary_DEG
summary_DEG_7
```


###############################################################################


Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Sid_Symbiont_DE_genes_OO_OC_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'OO-OC'
de.genes$group <- as.factor(de.genes$group)
dim(de.genes)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))

summary_DEG[1,3] <- summary_DEG[1,3]*(-1)

summary_DEG_8 <- summary_DEG
summary_DEG_8
```


###############################################################################

Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Sid_Symbiont_DE_genes_RO_RC_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'RO-RC'
de.genes$group <- as.factor(de.genes$group)
dim(de.genes)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))

summary_DEG[1,3] <- summary_DEG[1,3]*(-1)

summary_DEG_9 <- summary_DEG
summary_DEG_9
```


###############################################################################

# Effect of overall low pH vs. ambient pH
*However, no comparison of sites*

Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Sid_Symbiont_DE_genes_OO_RC_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'OO-RC'
de.genes$group <- as.factor(de.genes$group)
dim(de.genes)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))

summary_DEG[1,3] <- summary_DEG[1,3]*(-1)

summary_DEG_10 <- summary_DEG
summary_DEG_10
```


###############################################################################

# Effect of overall low pH vs. ambient pH
*However, no comparison of sites*

Second contrast

Absolute number of DEG adjusted p-value < 0.1
```{r}
# import data
de.genes <- readRDS("Sid_Symbiont_DE_genes_OO_LC_padj.1.rds")

# add group pairwise contrast label
de.genes$group <- 'OO-LC'
de.genes$group <- as.factor(de.genes$group)
dim(de.genes)
```

### Count which genes are Upregulated versus Downregulated
LFC = log2FoldChange
```{r}
de.genes$LFC <- ifelse(de.genes$log2FoldChange>0, 'Up', 'Down')
de.genes$LFC <- as.factor(de.genes$LFC)

summary_DEG <- de.genes %>% dplyr::count(group, LFC)
summary_DEG$LFC <- factor(summary_DEG$LFC, levels = c("Up", "Down"))

summary_DEG[1,3] <- summary_DEG[1,3]*(-1)

summary_DEG_11 <- summary_DEG
summary_DEG_11
```


###############################################################################


# DEG summary

```{r}
summary2 <- rbind(summary_DEG_1, summary_DEG_2, summary_DEG_3, summary_DEG_4, summary_DEG_5, summary_DEG_6, summary_DEG_7, summary_DEG_8, summary_DEG_9)

summary2
```


```{r}
summary2$Destination <- c("Control", "Control", "Control", "Control", "Control", "Control", "Ojo", "Ojo", "Ojo", "Ojo", "Ojo", "Ojo", "Ojo vs. Control", "Ojo vs. Control", "Ojo vs. Control", "Ojo vs. Control", "Ojo vs. Control", "Ojo vs. Control")

summary2$Destination <- as.factor(summary2$Destination)
summary2$Destination <- factor(summary2$Destination, levels = c("Control", "Ojo", "Ojo vs. Control"))

summary2
```


```{r}
# "OC-LC" "OC-RC" "LC-RC" "OO-LO" "LO-RO" "OO-RO" "LO-LC" "OO-OC" "RO-RC" "OO-RC" "OO-LC"
summary2$group <- factor(summary2$group, levels = c("OC-LC", "OC-RC", "LC-RC", "OO-LO", "OO-RO", "LO-RO", "LO-LC", "OO-OC", "RO-RC"))

write.csv(file = "DEG_summary_Sid_Symbiont_C.csv", x = summary2)
```


# Combined barplot
```{r}
plot.combo <- ggplot(summary2, aes(x = group, fill = LFC)) + 
  facet_wrap(~Destination, scales = "free_x") +
  geom_bar(data=subset(summary2, LFC=="Up"), aes(y=n, fill = LFC), stat = "identity", width=0.6) + #, binwidth=0
  geom_bar(data=subset(summary2, LFC=="Down"), aes(y=n, fill = LFC), stat = "identity", width=0.6) + #, binwidth=0
  geom_text(data=subset(summary2, LFC=="Up"), aes(group, n, label = n), position = position_dodge(width=0.9), vjust = -0.5) +
  geom_text(data=subset(summary2, LFC=="Down"), aes(group, n, label = -n), position = position_dodge(width=0.9), vjust = 1.5) +
  scale_fill_manual("LFC", values=c("black", "#56B4E9"), labels=c("Down-regulated", "Up-regulated"), guide=guide_legend(reverse=TRUE)) + #colorblind friendly
  ylab("Differentially expressed genes") +
  xlab("Transplantation") + 
  scale_y_continuous(limits = c(-5600, 500), breaks = seq(-5500, 500, by = 250)) +
  theme_classic() +
  theme(legend.position = "bottom", panel.background = element_rect(colour = "black"), axis.text.x = element_text(angle = 45, vjust = 0.5)) #axis.text.x=element_text(vjust=0.5, size=10, face="italic"), 

plot.combo
```

```{r}
# save plot
ggsave("Sid_Symbiont_C_DEG_barplot.pdf", plot = plot.combo,  width = 5, height = 4, units = "in", dpi = 1200) #width = 6.5, height = 4.5
```



## with low pH vs. ambient pH pairwise contrasts that *do not have comparative site contrast*

OO_RC
OO_LC

```{r}
summary3 <- rbind(summary_DEG_1, summary_DEG_2, summary_DEG_3, summary_DEG_4, summary_DEG_5, summary_DEG_6, summary_DEG_7, summary_DEG_8, summary_DEG_9, summary_DEG_10, summary_DEG_11)

summary3
```

```{r}
summary3$Destination <- c("Control", "Control", "Control", "Control", "Control", "Control", "Ojo", "Ojo", "Ojo", "Ojo", "Ojo", "Ojo", "Ojo vs. Control", "Ojo vs. Control", "Ojo vs. Control", "Ojo vs. Control", "Ojo vs. Control", "Ojo vs. Control", "Low vs. Ambient pH", "Low vs. Ambient pH", "Low vs. Ambient pH", "Low vs. Ambient pH")

summary3$Destination <- as.factor(summary3$Destination)
summary3$Destination <- factor(summary3$Destination, levels = c("Control", "Ojo", "Ojo vs. Control", "Low vs. Ambient pH"))

summary3
```


```{r}
# "OC-LC" "OC-RC" "LC-RC" "OO-LO" "LO-RO" "OO-RO" "LO-LC" "OO-OC" "RO-RC" "OO-RC" "OO-LC"
summary3$group <- factor(summary3$group, levels = c("OC-LC", "OC-RC", "LC-RC", "OO-LO", "OO-RO", "LO-RO", "LO-LC", "OO-OC", "RO-RC", "OO-RC", "OO-LC"))
```

```{r}
plot.combo <- ggplot(summary3, aes(x = group, fill = LFC)) + 
  facet_wrap(~Destination, scales = "free_x") +
  geom_bar(data=subset(summary3, LFC=="Up"), aes(y=n, fill = LFC), stat = "identity", width=0.6) + #, binwidth=0
  geom_bar(data=subset(summary3, LFC=="Down"), aes(y=n, fill = LFC), stat = "identity", width=0.6) + #, binwidth=0
  geom_text(data=subset(summary3, LFC=="Up"), aes(group, n, label = n), position = position_dodge(width=0.9), vjust = -0.5) +
  geom_text(data=subset(summary3, LFC=="Down"), aes(group, n, label = -n), position = position_dodge(width=0.9), vjust = 1.5) +
  scale_fill_manual("LFC", values=c("black", "#56B4E9"), labels=c("Down-regulated", "Up-regulated"), guide=guide_legend(reverse=TRUE)) + #colorblind friendly
  ylab("Differentially expressed genes") +
  xlab("Transplantation") + 
  scale_y_continuous(limits = c(-8200, 1200), breaks = seq(-7500, 1000, by = 500)) +
  theme_classic() +
  theme(legend.position = "bottom", panel.background = element_rect(colour = "black")) #axis.text.x=element_text(vjust=0.5, size=10, face="italic"), 

plot.combo

# save plot
#ggsave("Sid_Symbiont_C_DEG_barplot_4-dest-comparisons.pdf", plot = plot.combo,  width = 5, height = 4, units = "in", dpi = 1200) #width = 6.5, height = 4.5
```


### Session Info
```{r}
sessionInfo()
```



