---
title: "Venn_Sid_Sym_B"
author: "Veronica Radice"
date: "07/05/2021"
output: html_document
---


# Venn diagrams
## Differentially expressed genes

## Coral Symbionts - *Breviolum*
## *Siderastrea siderea*


### Load packages
```{r setup, include=FALSE}
library(VennDiagram)
library(limma)
library(GOplot)
library(RColorBrewer)
library(dplyr)
```

### Set working directory
```{r}
setwd("~/Documents/Rprojects/postdoc\ Rprojects/ODU_postdoc_Rprojects/Paper_Ojo_gene-expression/Siderastrea/symbiont/")
```


# Import data

Summary
```{r}
deg <- read.csv("DEG_summary_Sid_Symbiont_B.csv", header = TRUE)
deg <- deg[,2:4]
deg
```

## Effect of Origin - corals transplanted to same Destination

Absolute number of DEG adjusted p-value < 0.1

## Lagoon vs. Reef corals transplanted to Ojo
```{r}
de.genes.Dest.Ojo_LO_RO <- readRDS("Sid_Symbiont_Breviolum_DE_genes_LO_RO_padj.1.rds")
de.genes.Dest.Ojo_LO_RO <- as.data.frame(de.genes.Dest.Ojo_LO_RO)
#rownames(de.genes.Dest.Ojo_LO_RO) <- de.genes.Dest.Ojo_LO_RO$gene

de.genes.Dest.Ojo_LO_RO$group <- 'LO-RO'
de.genes.Dest.Ojo_LO_RO$group <- as.factor(de.genes.Dest.Ojo_LO_RO$group)

de.genes.Dest.Ojo_LO_RO$expression <- ifelse(de.genes.Dest.Ojo_LO_RO$log2FoldChange>0, 'up', 'down')
de.genes.Dest.Ojo_LO_RO$expression <- as.factor(de.genes.Dest.Ojo_LO_RO$expression)
dim(de.genes.Dest.Ojo_LO_RO)
```

```{r}
LO_RO.up <- subset(de.genes.Dest.Ojo_LO_RO, expression == "up")
dim(LO_RO.up)

LO_RO.down <- subset(de.genes.Dest.Ojo_LO_RO, expression == "down")
dim(LO_RO.down)
```


## Ojo vs. Reef corals transplanted to Ojo
```{r}
de.genes.Dest.Ojo_OO_RO <- readRDS("Sid_Symbiont_Breviolum_DE_genes_OO_RO_padj.1.rds")
de.genes.Dest.Ojo_OO_RO <- as.data.frame(de.genes.Dest.Ojo_OO_RO)
#rownames(de.genes.Dest.Ojo_OO_RO) <- de.genes.Dest.Ojo_OO_RO$gene

de.genes.Dest.Ojo_OO_RO$group <- 'OO-RO'
de.genes.Dest.Ojo_OO_RO$group <- as.factor(de.genes.Dest.Ojo_OO_RO$group)

de.genes.Dest.Ojo_OO_RO$expression <- ifelse(de.genes.Dest.Ojo_OO_RO$log2FoldChange>0, 'up', 'down')
de.genes.Dest.Ojo_OO_RO$expression <- as.factor(de.genes.Dest.Ojo_OO_RO$expression)
dim(de.genes.Dest.Ojo_OO_RO)
```

```{r}
OO_RO.up <- subset(de.genes.Dest.Ojo_OO_RO, expression == "up")
dim(OO_RO.up)

OO_RO.down <- subset(de.genes.Dest.Ojo_OO_RO, expression == "down")
dim(OO_RO.down)
```


## Lagoon vs. Reef corals transplanted to Control
```{r}
de.genes.Dest.Control_LC_RC <- readRDS("Sid_Symbiont_Breviolum_DE_genes_LC_RC_padj.1.rds")
de.genes.Dest.Control_LC_RC <- as.data.frame(de.genes.Dest.Control_LC_RC)
#rownames(de.genes.Dest.Control_LC_RC) <- de.genes.Dest.Control_LC_RC$gene

de.genes.Dest.Control_LC_RC$group <- 'LC-RC'
de.genes.Dest.Control_LC_RC$group <- as.factor(de.genes.Dest.Control_LC_RC$group)

de.genes.Dest.Control_LC_RC$expression <- ifelse(de.genes.Dest.Control_LC_RC$log2FoldChange>0, 'up', 'down')
de.genes.Dest.Control_LC_RC$expression <- as.factor(de.genes.Dest.Control_LC_RC$expression)
dim(de.genes.Dest.Control_LC_RC)
```

```{r}
LC_RC.up <- subset(de.genes.Dest.Control_LC_RC, expression == "up")
dim(LC_RC.up)

LC_RC.down <- subset(de.genes.Dest.Control_LC_RC, expression == "down")
dim(LC_RC.down)
```


## Ojo vs. Reef corals transplanted to Control
```{r}
de.genes.Dest.Control_OC_RC <- readRDS("Sid_Symbiont_Breviolum_DE_genes_OC_RC_padj.1.rds")
de.genes.Dest.Control_OC_RC <- as.data.frame(de.genes.Dest.Control_OC_RC)
#rownames(de.genes.Dest.Control_OC_RC) <- de.genes.Dest.Control_OC_RC$gene

de.genes.Dest.Control_OC_RC$group <- 'OC-RC'
de.genes.Dest.Control_OC_RC$group <- as.factor(de.genes.Dest.Control_OC_RC$group)

de.genes.Dest.Control_OC_RC$expression <- ifelse(de.genes.Dest.Control_OC_RC$log2FoldChange>0, 'up', 'down')
de.genes.Dest.Control_OC_RC$expression <- as.factor(de.genes.Dest.Control_OC_RC$expression)
dim(de.genes.Dest.Control_OC_RC)
```


```{r}
OC_RC.up <- subset(de.genes.Dest.Control_OC_RC, expression == "up")
dim(OC_RC.up)

OC_RC.down <- subset(de.genes.Dest.Control_OC_RC, expression == "down")
dim(OC_RC.down)
```


## Reef-Ojo vs. Reef-Control

```{r}
de.genes.Dest.Control_RO_RC <- readRDS("Sid_Symbiont_Breviolum_DE_genes_RO_RC_padj.1.rds")
de.genes.Dest.Control_RO_RC <- as.data.frame(de.genes.Dest.Control_RO_RC)
#rownames(de.genes.Dest.Control_RO_RC) <- de.genes.Dest.Control_RO_RC$gene

de.genes.Dest.Control_RO_RC$group <- 'RO-RC'
de.genes.Dest.Control_RO_RC$group <- as.factor(de.genes.Dest.Control_RO_RC$group)

de.genes.Dest.Control_RO_RC$expression <- ifelse(de.genes.Dest.Control_RO_RC$log2FoldChange>0, 'up', 'down')
de.genes.Dest.Control_RO_RC$expression <- as.factor(de.genes.Dest.Control_RO_RC$expression)
dim(de.genes.Dest.Control_RO_RC)
```

```{r}
RO_RC.up <- subset(de.genes.Dest.Control_RO_RC, expression == "up")
dim(RO_RC.up)

RO_RC.down <- subset(de.genes.Dest.Control_RO_RC, expression == "down")
dim(RO_RC.down)
```


```{r}
# data <- deg[1:2,2:3]
# data$n = abs(data$n)
# data <- data %>% rename(n = "OC-LC")

data <- deg[1:2,2:3]
data$n = abs(data$n)
data <- data %>% dplyr::rename('OC-RC' = n)
#data$`OC-RC` <- deg[3:4,3]
#data$`OC-RC` = abs(data$`OC-RC`)

data$`LC-RC` <- deg[3:4,3]
data$`LC-RC` = abs(data$`LC-RC`)

data$`LO-RO` <- deg[5:6,3]
data$`LO-RO` = abs(data$`LO-RO`)

data$`OO-RO` <- deg[7:8,3]
data$`OO-RO` = abs(data$`OO-RO`)

data$`RO-RC` <- deg[9:10,3]
data$`RO-RC` = abs(data$`RO-RC`)

data$LFC <- revalue(data$LFC, c("Down" = "down", "Up" = "up"))

rownames(data) <- data[,1]
data <- data[,-1]

data
```

# All DEG - 5-way venn comparing Control and Ojo destination pairwise contrasts
```{r}
venn.diagram(x = list(data$`LO-RO`, data$`OO-RO`, data$`LC-RC`, data$`OC-RC`, data$`RO-RC`), 
             category.names = c("Lagoon-Ojo vs Reef-Ojo", "Ojo-Ojo vs Reef-Ojo", "Lagoon-Control vs Reef-Control", "Ojo-Control vs Reef-Control", "Reef-Ojo vs Reef-Control"),
         filename = 'Sid_Symbiont_Breviolum_venn_5-way.png',
        output = TRUE,
        imagetype="png",
        height = 2500, 
        width = 4000, 
        resolution = 300,
        fill=c("cadetblue2", "darkseagreen2", "coral1", "deeppink3", "goldenrod"),
        main.cex = 1.5,
        sub.cex = 1.5,
        cat.cex = 1.5,
        cex = 1.5,
        cat.default.pos = "outer",
        cat.pos = c(-27, 27, 135, 135, 27),
        cat.dist = c(0.055, 0.055, 0.055, 0.055, 0.055),
        main = "Siderastrea Symbiont",
        scaled = FALSE)

#cat.pos = c(0, 0),
```


## Upregulated genes - with labels for reference
```{r}
venn.diagram(x = list(LO_RO.up$gene, OO_RO.up$gene, LC_RC.up$gene, OC_RC.up$gene, RO_RC.up$gene), 
             category.names = c("Lagoon-Ojo vs Reef-Ojo", "Ojo-Ojo vs Reef-Ojo", "Lagoon-Control vs Reef-Control", "Ojo-Control vs Reef-Control", "Reef-Ojo vs Reef-Control"),
         filename = 'Sid_Symbiont_Breviolum_venn_up_label.png',
        output = TRUE,
        imagetype="png",
        height = 2500, 
        width = 4000, 
        resolution = 300,
        fill=c("cadetblue2", "darkseagreen2", "coral1", "deeppink3", "goldenrod"),
        main.cex = 1.5,
        sub.cex = 1.5,
        cat.cex = 1.5,
        cex = 1.5,
        cat.default.pos = "outer",
        cat.pos = c(0, 0, 0, 0, 0), #c(0, 60, 150, 210, 350)
        cat.dist = c(0.055, 0.055, 0.055, 0.055, -0.1),
        main = "",
        scaled = FALSE)

#        cat.pos = c(-27, 27, 135, 135, 27),
#        cat.dist = c(0.055, 0.055, 0.055, 0.055, 0.055),
```


## Upregulated genes
```{r}
venn.diagram(x = list(LO_RO.up$gene, OO_RO.up$gene, LC_RC.up$gene, OC_RC.up$gene, RO_RC.up$gene), 
             category.names = c("", "", "", "", ""),
         filename = 'Sid_Symbiont_Breviolum_venn_up.png',
        output = TRUE,
        imagetype="png",
        height = 2500, 
        width = 4000, 
        resolution = 300,
        fill=c("cadetblue2", "darkseagreen2", "coral1", "deeppink3", "goldenrod"),
        main.cex = 1.5,
        sub.cex = 1.5,
        cat.cex = 1.5,
        cex = 1.5,
        cat.default.pos = "outer",
        cat.pos = c(0, 0, 0, 0, 0), #c(0, 60, 150, 210, 350)
        cat.dist = c(0.055, 0.055, 0.055, 0.055, -0.1),
        main = "",
        scaled = FALSE)

#        cat.pos = c(-27, 27, 135, 135, 27),
#        cat.dist = c(0.055, 0.055, 0.055, 0.055, 0.055),
```


## Downregulated genes
```{r}
venn.diagram(x = list(LO_RO.down$gene, OO_RO.down$gene, LC_RC.down$gene, OC_RC.down$gene, RO_RC.down$gene), 
             category.names = c("", "", "", "", ""),
         filename = 'Sid_Symbiont_Breviolum_venn_down.png',
        output = TRUE,
        imagetype="png",
        height = 2500, 
        width = 4000, 
        resolution = 300,
        fill=c("cadetblue2", "darkseagreen2", "coral1", "deeppink3", "goldenrod"),
        main.cex = 1.5,
        sub.cex = 1.5,
        cat.cex = 1.5,
        cex = 1.5,
        cat.default.pos = "outer",
        cat.pos = c(0, 0, 0, 0, 0), #c(0, 60, 150, 210, 350)
        cat.dist = c(0.055, 0.055, 0.055, 0.055, -0.1),
        main = "",
        scaled = FALSE)

#        cat.pos = c(-27, 27, 135, 135, 27),
#        cat.dist = c(0.055, 0.055, 0.055, 0.055, 0.055),
```

## Downregulated genes - with labels
```{r}
venn.diagram(x = list(LO_RO.down$gene, OO_RO.down$gene, LC_RC.down$gene, OC_RC.down$gene, RO_RC.down$gene), 
             category.names = c("Lagoon-Ojo vs Reef-Ojo", "Ojo-Ojo vs Reef-Ojo", "Lagoon-Control vs Reef-Control", "Ojo-Control vs Reef-Control", "Reef-Ojo vs Reef-Control"),
         filename = 'Sid_Symbiont_Breviolum_venn_down_label.png',
        output = TRUE,
        imagetype="png",
        height = 2500, 
        width = 4000, 
        resolution = 300,
        fill=c("cadetblue2", "darkseagreen2", "coral1", "deeppink3", "goldenrod"),
        main.cex = 1.5,
        sub.cex = 1.5,
        cat.cex = 1.5,
        cex = 1.5,
        cat.default.pos = "outer",
        cat.pos = c(0, 0, 0, 0, 0), #c(0, 60, 150, 210, 350)
        cat.dist = c(0.055, 0.055, 0.055, 0.055, -0.1),
        main = "",
        scaled = FALSE)

#        cat.pos = c(-27, 27, 135, 135, 27),
#        cat.dist = c(0.055, 0.055, 0.055, 0.055, 0.055),
```




### Session Info
```{r}
sessionInfo()
```

