---
title: "Venn_Sid_Host"
author: "Veronica Radice"
date: "19/04/2021"
output: html_document
---

# Venn diagrams
## Differentially expressed genes

## Coral Host
## *Siderastrea*


### Load packages
```{r setup, include=FALSE}
library(VennDiagram)
library(limma)
library(GOplot)
library(RColorBrewer)
```

### Set working directory
```{r}
setwd("~/Documents/Rprojects/postdoc\ Rprojects/ODU_postdoc_Rprojects/Paper_Ojo_gene-expression/Siderastrea/host/")
```


# Import data

Summary
```{r}
deg <- read.csv("DEG_summary_Sid_Host.csv", header = TRUE)
deg <- deg[,2:4]
deg
```

## Effect of Origin - corals transplanted to same Destination

Absolute number of DEG adjusted p-value < 0.1

## Lagoon vs. Reef corals transplanted to Ojo
```{r}
de.genes.Dest.Ojo_LO_RO <- readRDS("Sid_Host_DE_genes_LO_RO_padj.1.rds")
de.genes.Dest.Ojo_LO_RO <- as.data.frame(de.genes.Dest.Ojo_LO_RO)
#rownames(de.genes.Dest.Ojo_LO_RO) <- de.genes.Dest.Ojo_LO_RO$gene

de.genes.Dest.Ojo_LO_RO$group <- 'LO-RO'
de.genes.Dest.Ojo_LO_RO$group <- as.factor(de.genes.Dest.Ojo_LO_RO$group)

de.genes.Dest.Ojo_LO_RO$expression <- ifelse(de.genes.Dest.Ojo_LO_RO$log2FoldChange>0, 'up', 'down')
de.genes.Dest.Ojo_LO_RO$expression <- as.factor(de.genes.Dest.Ojo_LO_RO$expression)
dim(de.genes.Dest.Ojo_LO_RO)
```


## Ojo vs. Reef corals transplanted to Ojo
```{r}
de.genes.Dest.Ojo_OO_RO <- readRDS("Sid_Host_DE_genes_OO_RO_padj.1.rds")
de.genes.Dest.Ojo_OO_RO <- as.data.frame(de.genes.Dest.Ojo_OO_RO)
#rownames(de.genes.Dest.Ojo_OO_RO) <- de.genes.Dest.Ojo_OO_RO$gene

de.genes.Dest.Ojo_OO_RO$group <- 'OO-RO'
de.genes.Dest.Ojo_OO_RO$group <- as.factor(de.genes.Dest.Ojo_OO_RO$group)

de.genes.Dest.Ojo_OO_RO$expression <- ifelse(de.genes.Dest.Ojo_OO_RO$log2FoldChange>0, 'up', 'down')
de.genes.Dest.Ojo_OO_RO$expression <- as.factor(de.genes.Dest.Ojo_OO_RO$expression)
dim(de.genes.Dest.Ojo_OO_RO)
```


## Lagoon vs. Reef corals transplanted to Control
```{r}
de.genes.Dest.Control_LC_RC <- readRDS("Sid_Host_DE_genes_LC_RC_padj.1.rds")
de.genes.Dest.Control_LC_RC <- as.data.frame(de.genes.Dest.Control_LC_RC)
#rownames(de.genes.Dest.Control_LC_RC) <- de.genes.Dest.Control_LC_RC$gene

de.genes.Dest.Control_LC_RC$group <- 'LC-RC'
de.genes.Dest.Control_LC_RC$group <- as.factor(de.genes.Dest.Control_LC_RC$group)

de.genes.Dest.Control_LC_RC$expression <- ifelse(de.genes.Dest.Control_LC_RC$log2FoldChange>0, 'up', 'down')
de.genes.Dest.Control_LC_RC$expression <- as.factor(de.genes.Dest.Control_LC_RC$expression)
dim(de.genes.Dest.Control_LC_RC)
```


## Ojo vs. Reef corals transplanted to Control
```{r}
de.genes.Dest.Control_OC_RC <- readRDS("Sid_Host_DE_genes_OC_RC_padj.1.rds")
de.genes.Dest.Control_OC_RC <- as.data.frame(de.genes.Dest.Control_OC_RC)
#rownames(de.genes.Dest.Control_OC_RC) <- de.genes.Dest.Control_OC_RC$gene

de.genes.Dest.Control_OC_RC$group <- 'OC-RC'
de.genes.Dest.Control_OC_RC$group <- as.factor(de.genes.Dest.Control_OC_RC$group)

de.genes.Dest.Control_OC_RC$expression <- ifelse(de.genes.Dest.Control_OC_RC$log2FoldChange>0, 'up', 'down')
de.genes.Dest.Control_OC_RC$expression <- as.factor(de.genes.Dest.Control_OC_RC$expression)
dim(de.genes.Dest.Control_OC_RC)
```

## Reef-Ojo vs. Reef-Control

```{r}
de.genes.Dest.Control_RO_RC <- readRDS("Sid_Host_DE_genes_RO_RC_padj.1.rds")
de.genes.Dest.Control_RO_RC <- as.data.frame(de.genes.Dest.Control_RO_RC)
#rownames(de.genes.Dest.Control_RO_RC) <- de.genes.Dest.Control_RO_RC$gene

de.genes.Dest.Control_RO_RC$group <- 'RO-RC'
de.genes.Dest.Control_RO_RC$group <- as.factor(de.genes.Dest.Control_RO_RC$group)

de.genes.Dest.Control_RO_RC$expression <- ifelse(de.genes.Dest.Control_RO_RC$log2FoldChange>0, 'up', 'down')
de.genes.Dest.Control_RO_RC$expression <- as.factor(de.genes.Dest.Control_RO_RC$expression)
dim(de.genes.Dest.Control_RO_RC)
```


```{r}
# data <- deg[1:2,2:3]
# data$n = abs(data$n)
# data <- data %>% rename(n = "OC-LC")

data <- deg[3:4,2:3]
data$n = abs(data$n)
data <- data %>% rename(n = "OC-RC")
#data$`OC-RC` <- deg[3:4,3]
#data$`OC-RC` = abs(data$`OC-RC`)

data$`LC-RC` <- deg[5:6,3]
data$`LC-RC` = abs(data$`LC-RC`)

data$`LO-RO` <- deg[9:10,3]
data$`LO-RO` = abs(data$`LO-RO`)

data$`OO-RO` <- deg[11:12,3]
data$`OO-RO` = abs(data$`OO-RO`)

data$`RO-RC` <- deg[17:18,3]
data$`RO-RC` = abs(data$`RO-RC`)

data$LFC <- revalue(data$LFC, c("Down" = "down", "Up" = "up"))

rownames(data) <- data[,1]
data <- data[,-1]

data
```



```{r}
de.genes.Dest.Ojo_LO_RO %>% 
  group_by(expression) %>% 
  dplyr::summarise(count = n())
```


```{r}
deg <- rbind(de.genes.Dest.Ojo_LO_RO, de.genes.Dest.Ojo_OO_RO, de.genes.Dest.Control_LC_RC, de.genes.Dest.Control_OC_RC, de.genes.Dest.Control_RO_RC)
head(deg)
```

for GoVenn
Requires data frame consisting of two columns: ID, logFC
```{r}
de.genes.Dest.Ojo_LO_RO <- de.genes.Dest.Ojo_LO_RO[,c(1,3)]
names(de.genes.Dest.Ojo_LO_RO)[1] <- "ID"
names(de.genes.Dest.Ojo_LO_RO)[2] <- "logFC"

de.genes.Dest.Ojo_OO_RO <- de.genes.Dest.Ojo_OO_RO[,c(1,3)]
names(de.genes.Dest.Ojo_OO_RO)[1] <- "ID"
names(de.genes.Dest.Ojo_OO_RO)[2] <- "logFC"

de.genes.Dest.Control_LC_RC <- de.genes.Dest.Control_LC_RC[,c(1,3)]
names(de.genes.Dest.Control_LC_RC)[1] <- "ID"
names(de.genes.Dest.Control_LC_RC)[2] <- "logFC"

de.genes.Dest.Control_OC_RC <- de.genes.Dest.Control_OC_RC[,c(1,3)]
names(de.genes.Dest.Control_OC_RC)[1] <- "ID"
names(de.genes.Dest.Control_OC_RC)[2] <- "logFC"

de.genes.Dest.Control_RO_RC <- de.genes.Dest.Control_RO_RC[,c(1,3)]
names(de.genes.Dest.Control_RO_RC)[1] <- "ID"
names(de.genes.Dest.Control_RO_RC)[2] <- "logFC"
```




##############################################################################

# All DEG - 4-way venn comparing Control and Ojo destination pairwise contrasts
```{r}
venn.diagram(x = list(de.genes.Dest.Ojo_LO_RO$gene, de.genes.Dest.Ojo_OO_RO$gene, de.genes.Dest.Control_LC_RC$gene, de.genes.Dest.Control_OC_RC$gene, de.genes.Dest.Control_RO_RC$gene), 
             category.names = c("Lagoon-Ojo vs Reef-Ojo", "Ojo-Ojo vs Reef-Ojo", "Lagoon-Control vs Reef-Control", "Ojo-Control vs Reef-Control", "Reef-Ojo vs Reef-Control"),
         filename = 'Sid_Host_venn_all-DEG.png',
        output = TRUE,
        imagetype="png",
        height = 2500, 
        width = 4000, 
        resolution = 300,
        fill=c("cadetblue2", "darkseagreen2", "coral1", "deeppink3", "goldenrod"),
        main.cex = 1.5,
        sub.cex = 1.5,
        cat.cex = 1.5,
        cex = 1.5,
        cat.default.pos = "outer",
        cat.pos = c(-27, 27, 135, 135, 27),
        cat.dist = c(0.055, 0.055, 0.055, 0.055, 0.055),
        main = "Siderastrea",
        scaled = FALSE)

#cat.pos = c(0, 0),
```





Below is test of different Venn Diagram options


##############################################################################

## Using GoVenn from GoPlot

```{r}
GOVenn(de.genes.Dest.Ojo_LO_RO, de.genes.Dest.Ojo_OO_RO, de.genes.Dest.Control_LC_RC, de.genes.Dest.Control_OC_RC, de.genes.Dest.Control_RO_RC, lfc.col = c('cadetblue2', 'darkseagreen2', 'coral1', 'deeppink3', 'goldenrod'), label = c('Lagoon-Ojo vs Reef-Ojo', 'Ojo-Ojo vs Reef-Ojo', 'Lagoon-Control vs Reef-Control', 'Ojo-Control vs Reef-Control', 'Reef-Ojo vs Reef-Control'), plot = TRUE)
```

Can only do 3 contrasts at a time ??
```{r}
GOVenn(de.genes.Dest.Ojo_LO_RO, de.genes.Dest.Ojo_OO_RO, de.genes.Dest.Control_LC_RC, lfc.col = c('cadetblue2', 'darkseagreen2', 'coral1'), label = c('Lagoon-Ojo vs Reef-Ojo', 'Ojo-Ojo vs Reef-Ojo', 'Lagoon-Control vs Reef-Control'), plot = TRUE)
```




##############################################################################

# Using limma 
### include number of upregulated vs. downregulated DEG

```{r}
vennDiagram(data,
            include=c("up", "down"),
            counts.col=c("red", "blue"),
            circle.col = c("cadetblue2", "darkseagreen2", "coral1", "deeppink3", "goldenrod"))
```


```{r}
vennDiagram(deg,
            include=c("up", "down"),
            counts.col=c("red", "blue"),
            circle.col = c("red", "blue", "green3", "orange"))
```



##############################################################################

by lfc
```{r}
de.genes.Dest.Ojo_LO_RO <- de.genes.Dest.Ojo_LO_RO[,3]

de.genes.Dest.Ojo_OO_RO <- de.genes.Dest.Ojo_OO_RO[,3]

de.genes.Dest.Control_LC_RC <- de.genes.Dest.Control_LC_RC[,3]

de.genes.Dest.Control_OC_RC <- de.genes.Dest.Control_OC_RC[,3]

de.genes.Dest.Control_RO_RC <- de.genes.Dest.Control_RO_RC[,3]
```

by gene
```{r}
de.genes.Dest.Ojo_LO_RO <- de.genes.Dest.Ojo_LO_RO[,1]

de.genes.Dest.Ojo_OO_RO <- de.genes.Dest.Ojo_OO_RO[,1]

de.genes.Dest.Control_LC_RC <- de.genes.Dest.Control_LC_RC[,1]

de.genes.Dest.Control_OC_RC <- de.genes.Dest.Control_OC_RC[,1]

de.genes.Dest.Control_RO_RC <- de.genes.Dest.Control_RO_RC[,1]
```

```{r}
par(mar=c(0,0,0,0),mfrow=c(1,1),bty="n")
plot.new()
venn.plot<-venn.diagram(list(de.genes.Dest.Ojo_LO_RO, de.genes.Dest.Ojo_OO_RO, de.genes.Dest.Control_LC_RC, de.genes.Dest.Control_OC_RC, de.genes.Dest.Control_RO_RC), 
                        category.names = c("Lagoon-Ojo vs Reef-Ojo", "Ojo-Ojo vs Reef-Ojo", "Lagoon-Control vs Reef-Control", "Ojo-Control vs Reef-Control", "Reef-Ojo vs Reef-Control"),
                        height = 2500,
                        width = 4000,
                        filename = 'Sid_Host_venn_all-DEG_test.png',
                        output = TRUE,
                        fill=c("cadetblue2", "darkseagreen2", "coral1", "deeppink3", "goldenrod"), 
                        alpha=c(0.6, 0.6, 0.6, 0.6, 0.6), 
                        cex = 2, 
                        cat.fontface=4, 
                        main="Siderastrea DEG")
#grid.draw(venn.plot)
```





```{r}
SCt1<-read.csv("Spist_CBASS_29T1_vs_27T1_temp_and_gen_model.csv",header=T)
SCt2<-read.csv("Spist_CBASS_29T3_vs_27T3_temp_and_gen_model.csv",header=T)
SRt1<-read.csv("Spist_RSS_29T1_vs_27T1_temp_and_gen_model.csv",header=T)
SRt2<-read.csv("Spist_RSS_29T3_vs_27T3_temp_and_gen_model.csv",header=T)
SCt1<-SCt1$X
SCt2<-SCt2$X
SRt1<-SRt1$X
SRt2<-SRt2$X
par(mar=c(0,0,0,0),mfrow=c(1,1),bty="n")
plot.new()
venn.plot<-venn.diagram(list(SCt1 ,SRt2, SCt2,SRt1), height = 100, width = 300,NULL, fill=c("green", "blue", "yellow","red"), alpha=c(0.2, 0.2,0.2,0.2), cex = 2, cat.fontface=4, category.names=c("Spist CBASS T1", "Spist RSS T2","Spist CBASS T2","Spist RSS T1"), main="Gene DEG for 27 vs 29.5°C")
grid.draw(venn.plot)
```



```{r}
# sample five-set Venn Diagram
venn.plot <- venn.diagram(x = list(de.genes.Dest.Ojo_LO_RO, de.genes.Dest.Ojo_OO_RO, de.genes.Dest.Control_LC_RC, de.genes.Dest.Control_OC_RC, de.genes.Dest.Control_RO_RC),
                          category.names = c("Lagoon-Ojo vs Reef-Ojo", "Ojo-Ojo vs Reef-Ojo", "Lagoon-Control vs Reef-Control", "Ojo-Control vs Reef-Control", "Reef-Ojo vs Reef-Control"),
filename = "Venn_5set_pretty.tiff", col = "black",
fill = c("dodgerblue", "goldenrod1", "darkorange1", "seagreen3", "orchid3"),
alpha = 0.50,
#cex = c(1.5, 1.5, 1.5, 1.5, 1.5, 1, 0.8, 1, 0.8, 1, 0.8, 1, 0.8,
# 1, 0.8, 1, 0.55, 1, 0.55, 1, 0.55, 1, 0.55, 1, 0.55),
cat.col = c("dodgerblue", "goldenrod1", "darkorange1", "seagreen3", "orchid3"),
cat.cex = 1.5,
cat.fontface = "bold",
margin = 0.05)

venn.plot
```

Also
draw.quintuple.venn





### Session Info
```{r}
sessionInfo()
```

